\chapter{Robot Side API}

\label{AppendixRobotAPI}

\begin{longtable}{ l l p{8cm} }
\caption[Robot API]{The contents of the \textit{DebugNetwork} class that forms the robot side API.}\\
 \hline
 \multicolumn{3}{l}{\textbf{Variables}} \\
 \hline
 Name & Type & Purpose\\
 \hline\\
 socket\_ready & Boolean & Indicates whether the socket has been created successfully, and it ready to be used.\\
 sock\_in & Struct & A structure defining the target socket.\\
 sock\_fd & Integer & Identifies the socket on the local machine (robot).\\
 robot\_id & Integer & The numerical ID of this robot. Set at initialisation time. Inserted to the header of each packet to identify the sender.\\
 \hline
 \multicolumn{3}{l}{\textbf{Functions}} \\
 \hline
 Name & Return Type & Purpose\\
 \hline\\
 init & Void & Initialises the class by setting up the socket and setting the ID for this robot. Requires a port number, a target host IP address and the robot ID as arguments. \\
 destroy & Void & Shuts down the socket. \\
 sendData & Void & Sends a given string as a raw data packet. \\
 sendWatchdogPacket & Void & Constructs and sends a watchdog packet. Requires the robot name to be provided as an argument. \\
 sendStatePacket & Void & Constructs and sends a state packet. Requires the current state to be provided as an argument string. \\
 sendIRDataPacket & Void & Constructs and sends an IR data packet. Requires the data be arranged into an array, and takes a pointer to the first element and a size value as parameters. Also takes an extra boolean parameter which can be set to true to indicate that this packet should be identified as backgroud IR data.\\
 sendLogMessage & Void & Constructs and sends a message packet to display a message in the application console and logs. The message string is supplied as an argument.\\
 sendCustomData & Void & Constructs and sends a custom data packet. The key value pair are supplied to this function as two argument strings.\\
 getRobotID & Integer & Returns the numerical robot ID, as currently set. \\
 setRobotID & Void & Sets the numerical robot ID based on an integer argument. \\
 
 \label{tab:RobotAPI}
\end{longtable}